Worflow

### 2017-06-12 R/workflow.Rmd
Changes to preprocessin of the data

```{r}
runZinb <- T
runClus <- T
NCORES <- 7
```

@@

```{r parallel}
mysystem = Sys.info()[['sysname']]
switch(mysystem,
       Windows = {print("I'm a Windows PC and you should choose the parallel computing you want.")},
       Linux = {print("I'm a penguin and I'll use package MulticoreParam for parallel computing.")},
       Darwin = {print("I'm a Mac and I'll use package doParallel for parallel computing.")})

@@

```{r datain}
#counts<-read.table("../data/oeCufflinkCountData.txt",sep="\t",header=TRUE)
core <- read.table("../data/oeCufflinkCountData_1000Var.txt",
                   sep = "\t", header = TRUE)
core <- data.matrix(core)
metadata <- read.table("../data/oeMetadata.txt", sep = "\t", header = TRUE)
# symbol for samples missing from original clustering
metadata$clusterLabels[is.na(metadata$clusterLabels)] <- -2
```

@@

```{r batches}
batch <- metadata$Batch
```

@@

```{r qc}
qc <- metadata[, !names(metadata) %in% c("Batch", "Experiment", "clusterLabels")]
head(qc, 2)
```

```{r clusterlabels}
clus.labels <- metadata[, "clusterLabels"]
```

@@

```{r lookatdata}
dim(core)
core[1:3, 1:3]
```

Let's create a SummarizedExperiment object to store the raw counts and information about the data, that is batches, original labels, and quality control measures.

```{r se}
se <- SummarizedExperiment(assays = list(rawCounts = core),
                           colData = metadata)
```

@@

```{r zinb}
fn <- '../data/zinb_batch.rda'
if (runZinb & !file.exists(fn)){
  print(system.time(se <- zinbDimRed(se, K = 50, X = '~ Batch',
                                       residuals = T,
                                       normalizedValues = T)))
  save(se, file = fn)
}else{
  load(fn)
}
```

@@

```{r rsec_50}
fn <- '../data/RSEC_W.rda'
if (runClus & !file.exists(fn)){
  #symbol for samples missing from original clustering
	metadata$clusterLabels[is.na(metadata$clusterLabels)] <- -2 
  seObj <- SummarizedExperiment(t(W), colData = metadata)
  print(system.time(ceObj <- RSEC(seObj, k0s = 4:15, alphas = c(0.1),
                                  betas = 0.8,
                clusterFunction = "hierarchical01", minSizes=1,
                ncores = NCORES, isCount=FALSE,
                subsampleArgs = list(resamp.num=100,
                                     clusterFunction="kmeans",
                                     clusterArgs=list(nstart=10)),
                seqArgs = list(k.min=3, top.can=5), verbose=TRUE,
                combineProportion = 0.7,
                mergeMethod = "none")))
  save(ceObj, file = fn)
}else{
  load(fn)
}
```

@@

```{r refit_zinb}
fn <- '../data/refit_zinbwave_slingshot.rda'
if (runZinb & !file.exists(fn)){
  zinbList <- lapply(Kvec, function(k){
    zinbFit(se[, our_cl != "-1"], X = '~ Batch', K = k)
  })
  save(zinbList, file = fn)
}else{
  load(fn)
}
```

```{r slingshot_unsup_refit}
for(k in Kvec) {
  X <- getW(zinbList[[k - 2]])[, 1:k]

@@

```{r slingshot_sup_refit}
for(k in Kvec){
  X <- getW(zinbList[[k - 2]])[, 1:k]

@@

```{r de}
de <- read.csv('../data/oe_markers.txt', stringsAsFactors = F, header = F)
de <- de$V1
plotHeatmap(ceObj, 
            visualizeData = assays(se)$normalizedValues[rownames(se) %in% de, ],
            clusterSamplesData = "dendrogramValue",
            whichClusters = "primary",
            main = 'Normalized values, 1000 most variable genes',
            breaks = 0.99)
```

>>>>

```{r}
runZinb <- TRUE
runClus <- TRUE
NCORES <- 7
```

@@

```{r parallel}
mysystem = Sys.info()[['sysname']]
switch(mysystem,
       Windows = {print("I'm a Windows PC and I'll use the snow for parallel computing (warning: it might be slower than serial computations).")},
       Linux = {print("I'm a penguin and I'll use package MulticoreParam for parallel computing.")},
       Darwin = {print("I'm a Mac and I'll use package doParallel for parallel computing.")})

@@

DR: I have changed the `createData.R` file so that it uses scone to filter out low-quality samples.

```{r datain}
if(!file.exists("../data/oe_se_1000Var.rda")) {
  source("createData.R")
}

load("../data/oe_se_1000Var.rda")
```

@@

```{r batches}
batch <- colData(core)$Batch
```

@@

```{r qc}
qc <- colData(core)[, !names(colData(core)) %in% c("Batch", "Experiment", "clusterLabels")]
head(qc, 2)
```

```{r clusterlabels}
clus.labels <- colData(core)[, "clusterLabels"]
```

@@

```{r lookatdata}
dim(core)
assay(core)[1:3, 1:3]
```

@@

```{r zinb}
fn <- '../data/zinb_batch.rda'
if (runZinb & !file.exists(fn)){
  print(system.time(se <- zinbDimRed(core, K = 50, X = '~ Batch',
                                       residuals = TRUE,
                                       normalizedValues = TRUE)))
  save(se, file = fn)
}else{
  load(fn)
}
```

DR: the chunk above and the similar one for clusterExperiment are fine for now, but in the published workflow, we should just rely on the markdown cache system (the code above doesn't check for changes to the file or code -- just that a version of the file exists).

@@

```{r rsec_50}
fn <- '../data/RSEC_W.rda'
if (runClus & !file.exists(fn)){
  #symbol for samples missing from original clustering
  seObj <- SummarizedExperiment(t(W), colData = colData(core))
  print(system.time(ceObj <- RSEC(seObj, k0s = 4:15, alphas = c(0.1),
                                  betas = 0.8,
                clusterFunction = "hierarchical01", minSizes=1,
                ncores = NCORES, isCount=FALSE,
                subsampleArgs = list(resamp.num=100,
                                     clusterFunction="kmeans",
                                     clusterArgs=list(nstart=10)),
                seqArgs = list(k.min=3, top.can=5), verbose=TRUE,
                combineProportion = 0.7,
                mergeMethod = "none")))
  save(ceObj, file = fn)
}else{
  load(fn)
}
```

@@

```{r refit_zinb, eval=FALSE}
fn <- '../data/refit_zinbwave_slingshot.rda'

if (runZinb & !file.exists(fn)){
  zinbList <- lapply(Kvec, function(k){
    zinbFit(se[, our_cl != "-1"], X = '~ Batch', K = k)
  })
  save(zinbList, file = fn)
}else{
  load(fn)
}
```

```{r slingshot_unsup_refit, eval=FALSE}
for(k in Kvec) {
  X <- getW(zinbList[[k - 2]])[, 1:k]

@@

```{r slingshot_sup_refit, eval=FALSE}
for(k in Kvec){
  X <- getW(zinbList[[k - 2]])[, 1:k]

@@

```{r de, eval = FALSE}
de <- read.csv('../data/oe_markers.txt', stringsAsFactors = F, header = F)
de <- de$V1
plotHeatmap(ceObj, 
            visualizeData = assays(se)$normalizedValues[rownames(se) %in% de, ],
            clusterSamplesData = "dendrogramValue",
            whichClusters = "primary",
            main = 'Normalized values, 1000 most variable genes',
            breaks = 0.99)
```
